---
title: 'Week 4: Lab - Map Data and Visualization'
author: 'Brandon Eleff'
date: '2021-02-07'
output: html_document
---

Please see the discussion at https://thestorythusr.wordpress.com/2021/02/12/r-05-doing-math/

```{r setup, message = FALSE}

```

A demonstration of basic matrix operators and methods

```{r, "Basic Matrix Operations"}
# create some sample matrices
matrixA <- matrix(1:100, nrow = 10)
matrixB <- matrix(1:1000, nrow = 10)
matrixC <- matrix(c(1,2,3,4), nrow = 2)
matrixD <- matrix(c(5,6,7,8), nrow = 2)


# transpose
matrixA
t(matrixA)


# matrix multiplication -- commented out to make the output reasonable
# matrixA %*% matrixB

# inverse
solve(matrixC)

# determinant
det(matrixD)

# scalar and vector multiplication
2 * matrixA
vectorA <- c(5,6,7,8)
vectorA * matrixC

```


A simple example from game theory

In the game 2 players (the column player, and the row player) each pick a number from 1 to 4 (this game works for any range, by the way) and the player who picked the smaller number wins unless their number is exactly 1 less than the other players. If they pick the same number it is a draw.

We know that picking a lower number is good, so it would be silly to pick a 4. On the other hand the other player also knows picking lower numbers is good, so picking 1 all the time won't get us anywhere -- we need to mix it up a bit. We think it might be good to pick each number, 1-3 with a 1/3 probability, and we want to know what we can expect the payout to be long term...

Note that the game is symmetric and (1/3, 1/3, 1/3, 0) is in fact optimal for both players.
The upper left 3x3 if the matrix is exactly Rock-Paper-Scissors.

```{r, "A Simple Game"}
theGame<-matrix(c(0,1,-1,-1,-1,0,1,-1,1,-1,0,1,1,1,-1,0), nrow=4, byrow = TRUE)
theGame

randomStrat<-matrix( c(1/2,1/2,0,0), nrow = 4, byrow = TRUE)
MASS::fractions(randomStrat)

MASS::fractions(theGame %*% randomStrat)
```


A function to demonstrate an L2 regression
(note that R has this built in as lsfit())
```{r, "L2 Regression"}
#L2 Regression
# xBar = (aTranspose %*% a)^-1 %*% aTranspose %*% b

l2Regression <- function(matrixA, matrixB) {
aTranspose <- t(matrixA)
xBar <- solve(aTranspose %*% matrixA) %*% aTranspose %*% matrixB

return(xBar)
}

times <- c(1,2,3,5)
readings <- c(2,4,9,10)

timeMatrix <- matrix ( c(times,rep(1,length(times))), nrow = length(times))
readingsMatrix <- matrix(readings, nrow = length(readings))

# the solution is the slope and intercept of the line of best fit
# for least
solution<-l2Regression(timeMatrix, readingsMatrix)
MASS::fractions(solution)

plot(times,readings)
abline(solution[2], solution[1])
```



