---
title: 'Module 07: Objects'
author: 'usfeleff'
date: '2021-03-01'
output: html_document
---

Please see the discussion at https://thestorythusr.wordpress.com/2021/03/01/r-objects/

```{r setup, message = FALSE}
library(quantmod)
```
Download any type of data (from the web or use datasets package) or create your own set. 

Then, on the second step, determine if generic function as discussed in this module can be assigned to your data set, and if not, why? (Example, here is list of data set in R)
data("mtcars")
head (mtcars, 6)
list(mtcars, 6)

In third and last step, explore if S3 and S4 can be assigned to your data set.
 
In your blog, discuss the following questions:

How do you tell what OO system (S3 vs. S4) an object is associated with?

How do you determine the base type (like integer or list) of an object?

What is a generic function?

What are the main differences between S3 and S4?

In your GitHub, create two examples of S3 and S4.
```{r, "Datasets"}
# instance the built-in mtcars data set
myCars <- mtcars

# download stock data for apple (uses quantmod package), note the getSymbols()
# function creates a variable, no need for an assignment operator
getSymbols("AAPL", src="yahoo")

# download price of gold
getSymbols("XAU/USD",src="oanda")


```



```{r, "Generic Functions"}

# run the generic plot function against the 3 datasets, notice the difference
# in plot types
plot(myCars)
plot(myCars$mpg ~ myCars$hp)
plot(AAPL)
plot(XAUUSD)

```


```{r, "S3"}
# Create the list object and specify its class
s3Moto <- list(type="Motorcycle", make = "Honda", model_year= 2019, model = "CB300R", color = "black")
class(s3Moto) <- "motorcycle"


# view the object
s3Moto

# check the object type
mode(s3Moto)
isS4(s3Moto)
```

Notice that we cannot make an S4 object without first creating a class that
describes that object.

E.g., This line:
s4Moto <- new("Motorcycle", make = "Honda", model_year= 2019, model = "CB300R", color = "black")

Would fail with:
Error in getClass(Class, where = topenv(parent.frame())) : 
  “Motorcycle” is not a defined class

If we tried to run it before the setClasss()
```{r, "S4"}
# create an S4 (formal) Class for a Motorcycle
setClass("Motorcycle",
         representation(
           make ="character",
           model_year="numeric",
           model = "character",
           color = "character"
         )
         )

# create and view a particular motorcylce object
s4Moto <- new("Motorcycle", make = "Honda", model_year= 2019, model = "CB300R", color = "black")

s4Moto
# check the object type
mode(s4Moto)
isS4(s4Moto)
```
